Analysis-ready VCF at Biobank scale using Zarr

### 2025-01-31 paper.tex

[Discussion of the sparsity of the QC field data, and how this 
explains the difference to the VCF compression.]

>>>>

At 1.1X larger than the source gzipped VCF,
the relatively poor compression performance here 
is noteworthy and interesting. At only 7.44 GiB for
for 245K samples and 715K variants, VCF compression performance 
is exceptional and a result of several data reduction techniques.
Firstly, data for the AD, RGQ and FT fields is not stored 
for homozygous reference calls and the fields are consequently
extremely sparse. This is reflected in the Zarr, where 
compression performance for e.g.\ LAD is more than 10X higher
than the AD field in the Genomics England data
(Table~\ref{tab-genomics-england-data}).
Secondly, the GQ field has a very limited range of values 
as a result of gVCF reblocking. Together, these
techniques mean that the VCF text stream is highly compressible.
While the Zarr does not compress as well as VCF text in this 
case, the sparsity of the QC columns is reflected in the very
high compression rates. It is interesting to note that the 
BCF representation of this file is also larger than the VCF
(and Zarr) at 8.8GiB.
